### 2022-08-02 main/chapter_2ljets.tex
diboson to \diboson; clarify overlap removal

In the context of $\twoljets$ selections, $\metsig$ suppresses $\ttbar$
contributions more than $\met$ alone, plausibly since $\ttbar$ events are
accompanied by more hadronic activity than other sources of $\ptmiss$.
What remains of the standard model in the high-$\metsig$ tail of $\twoljets$
selections is almost pure diboson -- $ZZ$ with a hard $Z\rightarrow \nu\nu$,
and $W^\pm Z$ where the charged lepton in $W^\pm\rightarrow\ell^\pm\nu$
is not reconstructed.
\TODO{plot of diboson decomposition here?}
These effects are demonstrated from simulations in
Figure~\ref{fig:2ljets_presel_met}.

\begin{figure}[tp]
\centering
\includegraphics[width=0.48\textwidth]{figures/2ljets_presel_met_logy.png}
\hfill
\includegraphics[width=0.48\textwidth]{figures/2ljets_presel_met_sig_logy.png}
\caption[
Illustration of how $\metsig$ beats $\met$ in sensitivity to example signal
models
]{
Illustration of how $\metsig$ beats $\met$ in sensitivity to example signal
models.
At large $\metsig$, $\ttbar$ and other backgrounds are suppressed, leaving
quite pure diboson backgrounds and greater significance measures.
Signal models with smaller mass-splittings tend to have the bulk of their
events at smaller $\met$.
\\[0.5em]
This selection requires two same-flavour opposite-sign signal leptons with
$\pt^{\ell_2} > 25\,\eV[G]$,
$\mll \in (71, 111)\,\eV[G]$,
$\mjj \in (60, 110)\,\eV[G]$,
$\nbtag \leq 1$,
$\met > 100\,\eV[G]$, and
$\metsig > 6$.
\\[0.5em]
``Significance'' displayed in the lower subplot uses the Binomial significance
$Z_{Bi}$ from~\cite{cousins2008evaluation}, with a $30\%$ background
uncertainty and yields taken to the right of a given value.
Large values indicate expected sensitivity to the signal model.
}
\label{fig:2ljets_presel_met}
\end{figure}

@@

% presel table should come before High
\FloatBarrier
\subsection{High}
\label{sec:2ljets_high}
The highest $\metsig$ selections, of around $\metsig > 18$, contain almost pure
diboson backgrounds in relatively small quantities, along with the bulk masses
of signal samples at the heaviest resonance masses.
This is therefore the most promising territory for signal regions which are
highly sensitive to the most extreme signals.
Their backgrounds, however, must be constrained by control regions in less
extreme parameter spaces with more data.

@@

\begin{figure}[tp]
\centering
\begin{subfigure}{0.48\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/2ljets_region_design_hist1d_mll_SRHigh.png}
\caption{SR-High, $\mll$}
\end{subfigure}
\hfill
\begin{subfigure}{0.48\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/2ljets_region_design_hist1d_nbtag_SRHigh.png}
\caption{SR-High, $\nbtag$}
\end{subfigure}
\caption[
SR-High with relaxed selections on (a) $\mll$ or (b) $\nbtag$
]{%
SR-High with relaxed selections on (a) $\mll$ or (b) $\nbtag$.
The pure diboson background has very similar shapes to the signals.
Tighter cuts would therefore reduce background and signal yields comparably,
reducing sensitivity.
The significance dropouts in (a) are errors in plotting.
`Significance Z' is the binomial significance~\cite{cousins2008evaluation}
with a $30\%$ relative background uncertainty.
Error bars are statistical.
}
\label{fig:2ljets_high_mll_b}
\end{figure}

@@

One major difference between signals and diboson backgrounds is that signal
jet pairs mostly arise from boosted $W\rightarrow jj$ decays, which more often
produce nearby jets.
Selecting small $\rjj$ in SR-High therefore picks out these boosted cases.

@@

Intermediate signal regions are defined in Table~\ref{tab:2ljets_int},
along with their associated validation regions and the control regions in this
space.
To suppress the increased backgrounds, the intermediate $\mll$ window is
tightened to reduce non-$Z$ backgrounds, including $\ttbar$ and diboson
processes.
Requiring zero $b$-tags further reduces $\ttbar$ backgrounds.

@@

The control region \crvz\ occupies the $\mjj$ side-band to
SR-Int and VR-Int.
As seen from Figure~\ref{fig:2ljets_int_cr_region}, the selections in this
intermediate space give \crvz\ a very pure diboson background, which is very
valuable or constraining the modelling of High regions.
Also included in Table~\ref{tab:2ljets_int} and
Figure~\ref{fig:2ljets_int_cr_region} is \crtt, which uses various other
selections of $b$-tagged data to constrain $\ttbar$ backgrounds.
Both control regions show good agreement pre-fit and with statistical uncertainties
only.
\begin{figure}[tp]
\centering
\begin{subfigure}{0.48\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_CRZZ.png}
\caption{\crvz, $\metsig$}
\end{subfigure}
\hfill
\begin{subfigure}{0.48\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/2ljets_def_mjj_CRZZ.png}
\caption{\crvz, $\mjj$}
\end{subfigure}
\\[0.5em]
\begin{subfigure}{0.48\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_CRtt.png}
\caption{\crtt, $\metsig$}
\end{subfigure}
\hfill
\begin{subfigure}{0.48\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/2ljets_def_mjj_CRtt.png}
\caption{\crtt, $\mjj$}
\end{subfigure}
\caption[
Pre-fit distributions in the control regions CR-tt and CR-VZ
]{%
Pre-fit distributions in the control regions \crtt\ and \crvz.
Errors are statistical.
}
\label{fig:2ljets_int_cr_region}
\end{figure}

@@

Excluding $\metsig \in 9\textrm{--}12$ in SR-Low-2 is also not crucial.
This space contains about $2.3$ background expectation (mostly diboson) and
$\approx 1$ from some benchmark signals, so its inclusion as a single combined
bin not have increased sensitivity.
Its inclusion as an orthogonal bin with small yields would have added excessive
complication, however.

@@

The gap at in \crz\ at $\mjj \in 60\textrm{--}110\,\eV[G]$ is exactly SR-Low.
Their union therefore has a continuous $\mjj$ distribution, which is displayed
in Figure~\ref{fig:2ljets_low_sr_or_cr_region}.
Looking at the signal region alone can give an impression that $\zjets$ is
overestimated.
This plot contradicts that; most of the $\zjets$ distribution is fine;
the strange effect happens where diboson becomes large near $m(W)$.

@@

Additional control regions were designed in an attempt to improve background
normalization local to this region.
These are CR-Low-ISR-Z and CR-Low-ISR-DF, which select $\zjets$ and $\ttbar$
backgrounds and find smaller deficits of $20\%$ and $10\%$ respectively,
as shown in Figure~\ref{fig:2ljets_low_isr_old_more}.
To select $\zjets$, CR-Low-ISR-Z differs from CR-Low-ISR by requiring
$\dphiisrmet \in 1\textrm{--}2.6$ and $\metsig \in 6\textrm{--}9$.
To select $\ttbar$, CR-Low-ISR-DF differs from CR-Low-ISR by requiring
different-flavour leptons and the simpler $\mwisr > 20\,\eV[G]$, since it
has no signal regions to avoid in different-flavour data.
No nearby selection was found to form a pure diboson control region.

@@

A complicated likelihood model was constructed which normalized diboson,
$\zjets$ and $\ttbar$ backgrounds to the combination of CR-Low-ISR,
CR-Low-ISR-Z, CR-Low-ISR-DF.
After this fit, the validation region VR-Low-ISR had a high background
of $56.7 \pm 4.4$ for $47$ data, which is only a $1$-sigma deficit.

@@

Official jet mass scale (JMS) uncertainties for small-radius jets in the full
\atlas\ simulation were released late into the development of this search
(25 April 2021), and these uncertainties explicitly do not support the
fast simulation~\cite{atlas_twiki_smallrjms}.
These variations comprise six different parameters representing a reported
difference between data and simulations, differences between simulations from
different generators, uncertainties in tracking, and statistical uncertainties.
Once released, JMS variations were evaluated and used for the diboson and
$\ttbar$ contributions, which are the primary backgrounds in both 1J regions.

@@

To include other uncertainties that are not described by weights,
the standard approach is to compare the results of different simulation
algorithms, and we implement this for our major backgrounds $\zjets$, $\ttbar$,
and, to some extent diboson.
Each generator (simulation algorithm), or component thereof, uses different
approximate methods in its simulations, so the variations between simulations
indicate something about a reasonable uncertainty scale.

@@

Fitting only to control region (not signal regions) results in
Figure~\ref{fig:2ljets_low_crz_pre_post_pre}, which shows a clear deficit in
the signal region, with diboson yields peaking near to $m(W)$ above a smooth
varying $\zjets$ background.
Fitting to the signal region slashes the $\zjets$ background in SR-Low by
around $80\%$. How?
This sample comparison is the largest variation in VR-Low, and we have seen
from Table~\ref{tab:2ljets_zjets_alt} that SR-Low variations are around $30\%$.
A few factors contribute:
\begin{itemize}
\item The $\zjets$ sample is normalized to \crz\ with an unconstrained
normalization factor. This means that variations to the $\zjets$ sample in \crz
are absorbed by scaling all of its modifications in other regions.
\item As listed in Table~\ref{tab:2ljets_zjets_alt}, positive variations
in \crz\ match to negative variations in SR-Low bins; the two yields are
anticorrelated.
With normalization to \crz, this anticorrelation increases the effective
variation in SR-Low to $\pm40\%$.
\item Post-fit, the parameter controlling this $\zjets$ sample comparison
modifier is pulled to around $2\sigma$. This is seen in
Figure~\ref{fig:2ljets_fit_parameters}, and will be discussed further in
Section~\ref{sec:2ljets_results}.
\end{itemize}
A two-sigma pull on a $40\%$ scaling therefore explains this $80\%$ change.
Validation regions do not constrain our fits, and are also affected by
this effect; in VR-Low, small differences between the central and alternative
$\zjets$ samples are visible in Figure~\ref{fig:2ljets_vrlow_alt}, but
after fitting to the signal regions,
Figure~\ref{fig:2ljets_low_vrlow_pre_post} shows that the $\zjets$ sample
is almost wiped out, leaving a deficit below data.

\begin{figure}[tp]
\centering
\begin{subfigure}{0.48\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/2ljets_aux_CR_Z_EWKSR_Low_EWK_mjj_no_sr.png}
\caption{Fitted to control regions only}
\label{fig:2ljets_low_crz_pre_post_pre}
\end{subfigure}
\hfill
\begin{subfigure}{0.48\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/2ljets_aux_CR_Z_EWKSR_Low_EWK_mjj.pdf}
\caption{Fitted to control and signal regions}
\label{fig:2ljets_low_crz_pre_post_post}
\end{subfigure}
\caption[
Distribution of $\mjj$ in CR-Z and SR-Low with all systematic variations
]{%
Distribution of $\mjj$ in \crz\ and SR-Low with all systematic variations
(a) fitted to control regions, and (b) fitted to control and signal regions.
The arrows point in to the signal region, which shows a large peak in the
diboson background near $m(W)$.
Auxiliary materials to the paper~\cite{atlas2022searches} include (b), in which
the $\zjets$ background in SR-Low has been collapsed to fit the deficit.
Pre-fit yields are presented in Figures~\ref{fig:2ljets_low_sr_or_cr_region}
and~\ref{fig:2ljets_low_sr_or_cr_region_alt}.
}
\label{fig:2ljets_low_crz_pre_post}
\end{figure}

@@

To my eyes, Figure~\ref{fig:2ljets_low_crz_pre_post} suggests that the
excessive diboson background should also take some blame for this poor
prediction in SR-Low;
its diboson yields peak in the signal region on $\mjj \approx m(W)$, and
we do not expect large $W\rightarrow jj$ contributions in this diboson process,
since one boson is needed for the leptons and the other is needed for the
$\met$ that we select.
We shall see in Section~\ref{sec:2ljets_diboson_alt} why diboson yields are
not similarly squashed by sample comparison variations.

\subsubsection{Alternative diboson}
\label{sec:2ljets_diboson_alt}
No alternative samples were available for comparison with our central diboson
samples, which are simulated with \sherpa~2.2.2 at NLO with up to six
electroweak vertices.
This limitation is simply a result of how \atlas\ has decided to allocate
its limited resources on, and quite reasonably so, since full simulation is
computationally expensive.
We did, however, have access to the cheaper alternative set of small,
truth-level samples, that were simulated with the same \sherpa\ setup but with
varied parameters to assess uncertainties in the modelling of parton showers.

These small truth samples include variations of three parameters:
\begin{itemize}
% CKKW: merging scale
\item $Q_\textrm{cut}^\textrm{CKKW}$, the energy scale for merging
matrix elements with parton showers under the CKKW
scheme~\cite{Catani_2001_CKKW,Hoeche:2009rj}.
It is varied from its default $20\,\eV[G]$ up to $30\,\eV[G]$ and down
to $15\,\eV[G]$.
% CSS_KIN: CSS is the combined matrix element + shower
\item \texttt{CSS\_KIN}, a binary parameter that tells \sherpa\ how to treat
kinematic recoils in its parton showers~\cite{sherpa_222_manual}.
It has two modes, one default~\cite{Hoeche:2009rj}, and one
alternative~\cite{Hoeche:2009xc}
that is recommended for studies of systematic variations.
These differ in how they resolve the problem that an isolated massless parton
cannot conserve four-momentum when splitting into two non-collinear massless
children (for example gluon emission $q\rightarrow gq$);
each scheme achieves conservation by recoiling off different parts of the
parton shower.
% resummation: parton showing including all orders in QCD
% https://link.springer.com/article/10.1007/JHEP01(2018)118
% "radiative corrections determined in QCD perturbation theory must be resummed
% to all orders."
% https://sherpa-team.gitlab.io/sherpa/tmp-update-extamp-examples/manual/parameters/matrix-elements.html
% "In parton shower matched/merged calculations a third perturbative scale is
% present, the resummation or parton shower starting scale. It can be set by
% the user in the third argument like"
% METS is a scale setter
% QSF: scale factor for the resummation scale; relates to parton showers; quadratic
% vary up and down by a factor of 4, but QSF is quadratic in the scale so this is
% an implementation details of the usual 2x 0.5x form
\item $S_Q$, a scale factor on the resummation scale $\mu_{Q^2}$ that is set
from the hard scatter kinematics in each event and affects emissions in the
parton shower.
Its central value is $1$, and it is varied up and down by factors of $4$.
Since \sherpa\ defines its scales in a quadratic form, this matches the typical
convention of considering scale variations over factors of
$2$~\cite{sherpa_222_manual,Bothmann:2019yzt}.
\end{itemize}
Truth-level background yields in our regions were calculated with the
SimpleAnalysis~\cite{Lorenz:2752450,simpleanalysis_cern}
software package for
our central diboson samples and these varied alternatives.
SimpleAnalysis histograms truth-level simulations and optionally smears them
with calibrated noise to approximate reconstruction-level results.
\emph{Thanks to Jon~Long for preparing our initial
SimpleAnalysis implementations and sharing software to evaluate these
variations.}

@@

This assignment varies \crvz\ and on-shell other regions identically,
and since the diboson background is normalized to \crvz\ this means that
these variations have no effect in on-shell regions, and cannot move to
fit the SR-Low or SR-High-8 deficits.

Had we considered complete alternative diboson samples, we would certainly
have had some meaningful variations to apply in those regions, but I do not
know how large those variations should be.
This lack of a real alternative sample is particularly disappointing since
diboson is our most important background.
In hindsight, it may have been sensible to push for more resources to be
allocated to resolve this problem.

@@

Since these variations are highly correlated between regions, samples that are
normalized to control regions (diboson, $\ttbar$, $\zjets$, and $\zyjets$)
absorb some of their effect into their unconstrained normalization factors.
For other samples, the effect of scale variations is already included in
the uncertainty on their total cross-section, so their scale variations are
normalized to the sum of varied weights in very loose pre-selections,
similarly reducing their effects.

@@

\paragraph{Miscellanea}
\begin{itemize}
\item A $\pm30\%$ variation is applied to $\zjets$ in \crtt\ to cover our
uncertainty from extrapolation over $b$-tagged
jets~\cite{atlas_twiki_susytheoretical}.
\item Variations for the \sherpa\ parton shower parameters
$Q_\textrm{cut}^\textrm{CKKW}$ and
$S_Q$ are assigned for $\zjets$.
Unlike the alternative samples used for diboson as described in
Section~\ref{sec:2ljets_diboson_alt}, these are assigned as functions of the
jet kinematics in each event by a parametrized method that avoids expensive
alternative simulations~\cite{Anders:2125718, atlas_twiki_susytheoretical}.
\emph{Thanks to Jon~Long and Yumeng~Cao for collaboration in preparing these
variations}.
\end{itemize}

@@

Yet another unique sample-comparison prescription now assigns variations as
half of the absolute difference between up and down yields, irrespective of the
central yield.
This differs from the methods previously described in
Sections~\ref{sec:2ljets_zjets_alt} and~\ref{sec:2ljets_diboson_alt},
and is used for consistency with other $\twoljets$ work.

@@

\paragraph{Sub-processes}
Our high-level sample labels such as $\ttbar$, $\zjets$, diboson and
rare-top (`Top other') each include many sub-samples that are physically
distinct and non-interfering.
Highlights from a study of the contributions from these sub-samples are
presented in Figure~\ref{fig:2ljets_splits}.

>>>>

In the context of $\twoljets$ selections, $\metsig$ suppresses $\ttbar$
contributions more than $\met$ alone, plausibly since $\ttbar$ events are
accompanied by more hadronic activity than other sources of $\ptmiss$.
What remains of the standard model in the high-$\metsig$ tail of $\twoljets$
selections is almost pure \diboson\ -- $ZZ$ with a hard $Z\rightarrow \nu\nu$,
and $W^\pm Z$ where the charged lepton in $W^\pm\rightarrow\ell^\pm\nu$
is not reconstructed.
\TODO{plot of \diboson\ decomposition here?}
These effects are demonstrated from simulations in
Figure~\ref{fig:2ljets_presel_met}.

\begin{figure}[tp]
\centering
\includegraphics[width=0.48\textwidth]{figures/2ljets_presel_met_logy.png}
\hfill
\includegraphics[width=0.48\textwidth]{figures/2ljets_presel_met_sig_logy.png}
\caption[
Illustration of how $\metsig$ beats $\met$ in sensitivity to example signal
models
]{
Illustration of how $\metsig$ beats $\met$ in sensitivity to example signal
models.
At large $\metsig$, $\ttbar$ and other backgrounds are suppressed, leaving
quite pure \diboson\ backgrounds and greater significance measures.
Signal models with smaller mass-splittings tend to have the bulk of their
events at smaller $\met$.
\\[0.5em]
This selection requires two same-flavour opposite-sign signal leptons with
$\pt^{\ell_2} > 25\,\eV[G]$,
$\mll \in (71, 111)\,\eV[G]$,
$\mjj \in (60, 110)\,\eV[G]$,
$\nbtag \leq 1$,
$\met > 100\,\eV[G]$, and
$\metsig > 6$.
\\[0.5em]
``Significance'' displayed in the lower subplot uses the Binomial significance
$Z_{Bi}$ from~\cite{cousins2008evaluation}, with a $30\%$ background
uncertainty and yields taken to the right of a given value.
Large values indicate expected sensitivity to the signal model.
}
\label{fig:2ljets_presel_met}
\end{figure}

@@

% presel table should come before High
\FloatBarrier
\subsection{High}
\label{sec:2ljets_high}
The highest $\metsig$ selections, of around $\metsig > 18$, contain almost pure
\diboson\ backgrounds in relatively small quantities, along with the bulk masses
of signal samples at the heaviest resonance masses.
This is therefore the most promising territory for signal regions which are
highly sensitive to the most extreme signals.
Their backgrounds, however, must be constrained by control regions in less
extreme parameter spaces with more data.

@@

\begin{figure}[tp]
\centering
\begin{subfigure}{0.48\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/2ljets_region_design_hist1d_mll_SRHigh.png}
\caption{SR-High, $\mll$}
\end{subfigure}
\hfill
\begin{subfigure}{0.48\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/2ljets_region_design_hist1d_nbtag_SRHigh.png}
\caption{SR-High, $\nbtag$}
\end{subfigure}
\caption[
SR-High with relaxed selections on (a) $\mll$ or (b) $\nbtag$
]{%
SR-High with relaxed selections on (a) $\mll$ or (b) $\nbtag$.
The pure \diboson\ background has very similar shapes to the signals.
Tighter cuts would therefore reduce background and signal yields comparably,
reducing sensitivity.
The significance dropouts in (a) are errors in plotting.
`Significance Z' is the binomial significance~\cite{cousins2008evaluation}
with a $30\%$ relative background uncertainty.
Error bars are statistical.
}
\label{fig:2ljets_high_mll_b}
\end{figure}

@@

One major difference between signals and \diboson\ backgrounds is that signal
jet pairs mostly arise from boosted $W\rightarrow jj$ decays, which more often
produce nearby jets.
Selecting small $\rjj$ in SR-High therefore picks out these boosted cases.

@@

Intermediate signal regions are defined in Table~\ref{tab:2ljets_int},
along with their associated validation regions and the control regions in this
space.
To suppress the increased backgrounds, the intermediate $\mll$ window is
tightened to reduce non-$Z$ backgrounds, including $\ttbar$ and \diboson
processes.
Requiring zero $b$-tags further reduces $\ttbar$ backgrounds.

@@

The control region \crvz\ occupies the $\mjj$ side-band to
SR-Int and VR-Int.
As seen from Figure~\ref{fig:2ljets_int_cr_region}, the selections in this
intermediate space give \crvz\ a very pure \diboson\ background, which is very
valuable or constraining the modelling of High regions.
Also included in Table~\ref{tab:2ljets_int} and
Figure~\ref{fig:2ljets_int_cr_region} is \crtt, which uses various other
selections of $b$-tagged data to constrain $\ttbar$ backgrounds.
Both control regions show good agreement pre-fit and with statistical uncertainties
only.
\begin{figure}[tp]
\centering
\begin{subfigure}{0.48\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_CRZZ.png}
\caption{\crvz, $\metsig$}
\end{subfigure}
\hfill
\begin{subfigure}{0.48\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/2ljets_def_mjj_CRZZ.png}
\caption{\crvz, $\mjj$}
\end{subfigure}
\\[0.5em]
\begin{subfigure}{0.48\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/2ljets_def_met_Sign_CRtt.png}
\caption{\crtt, $\metsig$}
\end{subfigure}
\hfill
\begin{subfigure}{0.48\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/2ljets_def_mjj_CRtt.png}
\caption{\crtt, $\mjj$}
\end{subfigure}
\caption[
Pre-fit distributions in the control regions CR-tt and CR-VZ
]{%
Pre-fit distributions in the control regions \crtt\ and \crvz.
Errors are statistical.
}
\label{fig:2ljets_int_cr_region}
\end{figure}

@@

Excluding $\metsig \in 9\textrm{--}12$ in SR-Low-2 is also not crucial.
This space contains about $2.3$ background expectation (mostly \diboson) and
$\approx 1$ from some benchmark signals, so its inclusion as a single combined
bin not have increased sensitivity.
Its inclusion as an orthogonal bin with small yields would have added excessive
complication, however.

@@

The gap at in \crz\ at $\mjj \in 60\textrm{--}110\,\eV[G]$ is exactly SR-Low.
Their union therefore has a continuous $\mjj$ distribution, which is displayed
in Figure~\ref{fig:2ljets_low_sr_or_cr_region}.
Looking at the signal region alone can give an impression that $\zjets$ is
overestimated.
This plot contradicts that; most of the $\zjets$ distribution is fine;
the strange effect happens where \diboson\ becomes large near $m(W)$.

@@

Additional control regions were designed in an attempt to improve background
normalization local to this region.
These are CR-Low-ISR-Z and CR-Low-ISR-DF, which select $\zjets$ and $\ttbar$
backgrounds and find smaller deficits of $20\%$ and $10\%$ respectively,
as shown in Figure~\ref{fig:2ljets_low_isr_old_more}.
To select $\zjets$, CR-Low-ISR-Z differs from CR-Low-ISR by requiring
$\dphiisrmet \in 1\textrm{--}2.6$ and $\metsig \in 6\textrm{--}9$.
To select $\ttbar$, CR-Low-ISR-DF differs from CR-Low-ISR by requiring
different-flavour leptons and the simpler $\mwisr > 20\,\eV[G]$, since it
has no signal regions to avoid in different-flavour data.
No nearby selection was found to form a pure \diboson\ control region.

@@

A complicated likelihood model was constructed which normalized \diboson,
$\zjets$ and $\ttbar$ backgrounds to the combination of CR-Low-ISR,
CR-Low-ISR-Z, CR-Low-ISR-DF.
After this fit, the validation region VR-Low-ISR had a high background
of $56.7 \pm 4.4$ for $47$ data, which is only a $1$-sigma deficit.

@@

Official jet mass scale (JMS) uncertainties for small-radius jets in the full
\atlas\ simulation were released late into the development of this search
(25 April 2021), and these uncertainties explicitly do not support the
fast simulation~\cite{atlas_twiki_smallrjms}.
These variations comprise six different parameters representing a reported
difference between data and simulations, differences between simulations from
different generators, uncertainties in tracking, and statistical uncertainties.
Once released, JMS variations were evaluated and used for the \diboson\ and
$\ttbar$ contributions, which are the primary backgrounds in both 1J regions.

@@

To include other uncertainties that are not described by weights,
the standard approach is to compare the results of different simulation
algorithms, and we implement this for our major backgrounds $\zjets$, $\ttbar$,
and, to some extent \diboson.
Each generator (simulation algorithm), or component thereof, uses different
approximate methods in its simulations, so the variations between simulations
indicate something about a reasonable uncertainty scale.

@@

Fitting only to control region (not signal regions) results in
Figure~\ref{fig:2ljets_low_crz_pre_post_pre}, which shows a clear deficit in
the signal region, with \diboson\ yields peaking near to $m(W)$ above a smooth
varying $\zjets$ background.
Fitting to the signal region slashes the $\zjets$ background in SR-Low by
around $80\%$. How?
This sample comparison is the largest variation in VR-Low, and we have seen
from Table~\ref{tab:2ljets_zjets_alt} that SR-Low variations are around $30\%$.
A few factors contribute:
\begin{itemize}
\item The $\zjets$ sample is normalized to \crz\ with an unconstrained
normalization factor. This means that variations to the $\zjets$ sample in \crz
are absorbed by scaling all of its modifications in other regions.
\item As listed in Table~\ref{tab:2ljets_zjets_alt}, positive variations
in \crz\ match to negative variations in SR-Low bins; the two yields are
anticorrelated.
With normalization to \crz, this anticorrelation increases the effective
variation in SR-Low to $\pm40\%$.
\item Post-fit, the parameter controlling this $\zjets$ sample comparison
modifier is pulled to around $2\sigma$. This is seen in
Figure~\ref{fig:2ljets_fit_parameters}, and will be discussed further in
Section~\ref{sec:2ljets_results}.
\end{itemize}
A two-sigma pull on a $40\%$ scaling therefore explains this $80\%$ change.
Validation regions do not constrain our fits, and are also affected by
this effect; in VR-Low, small differences between the central and alternative
$\zjets$ samples are visible in Figure~\ref{fig:2ljets_vrlow_alt}, but
after fitting to the signal regions,
Figure~\ref{fig:2ljets_low_vrlow_pre_post} shows that the $\zjets$ sample
is almost wiped out, leaving a deficit below data.

\begin{figure}[tp]
\centering
\begin{subfigure}{0.48\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/2ljets_aux_CR_Z_EWKSR_Low_EWK_mjj_no_sr.png}
\caption{Fitted to control regions only}
\label{fig:2ljets_low_crz_pre_post_pre}
\end{subfigure}
\hfill
\begin{subfigure}{0.48\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/2ljets_aux_CR_Z_EWKSR_Low_EWK_mjj.pdf}
\caption{Fitted to control and signal regions}
\label{fig:2ljets_low_crz_pre_post_post}
\end{subfigure}
\caption[
Distribution of $\mjj$ in CR-Z and SR-Low with all systematic variations
]{%
Distribution of $\mjj$ in \crz\ and SR-Low with all systematic variations
(a) fitted to control regions, and (b) fitted to control and signal regions.
The arrows point in to the signal region, which shows a large peak in the
\diboson\ background near $m(W)$.
Auxiliary materials to the paper~\cite{atlas2022searches} include (b), in which
the $\zjets$ background in SR-Low has been collapsed to fit the deficit.
Pre-fit yields are presented in Figures~\ref{fig:2ljets_low_sr_or_cr_region}
and~\ref{fig:2ljets_low_sr_or_cr_region_alt}.
}
\label{fig:2ljets_low_crz_pre_post}
\end{figure}

@@

To my eyes, Figure~\ref{fig:2ljets_low_crz_pre_post} suggests that the
excessive \diboson\ background should also take some blame for this poor
prediction in SR-Low;
its \diboson\ yields peak in the signal region on $\mjj \approx m(W)$, and
we do not expect large $W\rightarrow jj$ contributions in this \diboson\ process,
since one boson is needed for the leptons and the other is needed for the
$\met$ that we select.
We shall see in Section~\ref{sec:2ljets_\diboson_alt} why \diboson\ yields are
not similarly squashed by sample comparison variations.

\subsubsection{Alternative \diboson}
\label{sec:2ljets_\diboson_alt}
No alternative samples were available for comparison with our central \diboson
samples, which are simulated with \sherpa~2.2.2 at NLO with up to six
electroweak vertices.
This limitation is simply a result of how \atlas\ has decided to allocate
its limited resources on, and quite reasonably so, since full simulation is
computationally expensive.
We did, however, have access to the cheaper alternative set of small,
truth-level samples, that were simulated with the same \sherpa\ setup but with
varied parameters to assess uncertainties in the modelling of parton showers.

These small truth samples include variations of three parameters:
\begin{itemize}
% CKKW: merging scale
\item $Q_\textrm{cut}^\textrm{CKKW}$, the energy scale for merging
matrix elements with parton showers under the CKKW
scheme~\cite{Catani_2001_CKKW,Hoeche:2009rj}.
It is varied from its default $20\,\eV[G]$ up to $30\,\eV[G]$ and down
to $15\,\eV[G]$.
% CSS_KIN: CSS is the combined matrix element + shower
\item \texttt{CSS\_KIN}, a binary parameter that tells \sherpa\ how to treat
kinematic recoils in its parton showers~\cite{sherpa_222_manual}.
It has two modes, one default~\cite{Hoeche:2009rj}, and one
alternative~\cite{Hoeche:2009xc}
that is recommended for studies of systematic variations.
These differ in how they resolve the problem that an isolated massless parton
cannot conserve four-momentum when splitting into two non-collinear massless
children (for example gluon emission $q\rightarrow gq$);
each scheme achieves conservation by recoiling off different parts of the
parton shower.
% resummation: parton showing including all orders in QCD
% https://link.springer.com/article/10.1007/JHEP01(2018)118
% "radiative corrections determined in QCD perturbation theory must be resummed
% to all orders."
% https://sherpa-team.gitlab.io/sherpa/tmp-update-extamp-examples/manual/parameters/matrix-elements.html
% "In parton shower matched/merged calculations a third perturbative scale is
% present, the resummation or parton shower starting scale. It can be set by
% the user in the third argument like"
% METS is a scale setter
% QSF: scale factor for the resummation scale; relates to parton showers; quadratic
% vary up and down by a factor of 4, but QSF is quadratic in the scale so this is
% an implementation details of the usual 2x 0.5x form
\item $S_Q$, a scale factor on the resummation scale $\mu_{Q^2}$ that is set
from the hard scatter kinematics in each event and affects emissions in the
parton shower.
Its central value is $1$, and it is varied up and down by factors of $4$.
Since \sherpa\ defines its scales in a quadratic form, this matches the typical
convention of considering scale variations over factors of
$2$~\cite{sherpa_222_manual,Bothmann:2019yzt}.
\end{itemize}
Truth-level background yields in our regions were calculated with the
SimpleAnalysis~\cite{Lorenz:2752450,simpleanalysis_cern}
software package for
our central \diboson\ samples and these varied alternatives.
SimpleAnalysis histograms truth-level simulations and optionally smears them
with calibrated noise to approximate reconstruction-level results.
\emph{Thanks to Jon~Long for preparing our initial
SimpleAnalysis implementations and sharing software to evaluate these
variations.}

@@

This assignment varies \crvz\ and on-shell other regions identically,
and since the \diboson\ background is normalized to \crvz\ this means that
these variations have no effect in on-shell regions, and cannot move to
fit the SR-Low or SR-High-8 deficits.

Had we considered complete alternative \diboson\ samples, we would certainly
have had some meaningful variations to apply in those regions, but I do not
know how large those variations should be.
This lack of a real alternative sample is particularly disappointing since
\diboson\ is our most important background.
In hindsight, it may have been sensible to push for more resources to be
allocated to resolve this problem.

@@

Since these variations are highly correlated between regions, samples that are
normalized to control regions (\diboson, $\ttbar$, $\zjets$, and $\zyjets$)
absorb some of their effect into their unconstrained normalization factors.
For other samples, the effect of scale variations is already included in
the uncertainty on their total cross-section, so their scale variations are
normalized to the sum of varied weights in very loose pre-selections,
similarly reducing their effects.

@@

\paragraph{Miscellanea}
\begin{itemize}
\item A $\pm30\%$ variation is applied to $\zjets$ in \crtt\ to cover our
uncertainty from extrapolation over $b$-tagged
jets~\cite{atlas_twiki_susytheoretical}.
\item Variations for the \sherpa\ parton shower parameters
$Q_\textrm{cut}^\textrm{CKKW}$ and
$S_Q$ are assigned for $\zjets$.
Unlike the alternative samples used for \diboson\ as described in
Section~\ref{sec:2ljets_\diboson_alt}, these are assigned as functions of the
jet kinematics in each event by a parametrized method that avoids expensive
alternative simulations~\cite{Anders:2125718, atlas_twiki_susytheoretical}.
\emph{Thanks to Jon~Long and Yumeng~Cao for collaboration in preparing these
variations}.
\end{itemize}

@@

Yet another unique sample-comparison prescription now assigns variations as
half of the absolute difference between up and down yields, irrespective of the
central yield.
This differs from the methods previously described in
Sections~\ref{sec:2ljets_zjets_alt} and~\ref{sec:2ljets_\diboson_alt},
and is used for consistency with other $\twoljets$ work.

@@

\paragraph{Sub-processes}
Our high-level sample labels such as $\ttbar$, $\zjets$, \diboson\ and
rare-top (`Top other') each include many sub-samples that are physically
distinct and non-interfering.
Highlights from a study of the contributions from these sub-samples are
presented in Figure~\ref{fig:2ljets_splits}.
